Numerical investigation of viscous flow instabilities through heterogeneous porous media. Advantages and limitations of unstructured and adaptive meshes in 2D and 3D domains under different mobility scenarios.

### 2017-12-07 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
change group name

In this section, 3D simulations are introduced to investigate preferential flow pathways and finger growth through semi-pervious and impervious layers, as decribed in Figure \ref{fig:3DChannel_PermMesh}. In large scale geological formations preferential flow may allow fast contaminant transport (therefore creating significant consequences for groundwater quality) and low recovery rate (in hydrocarbon production fields).

As mentioned in Section \ref{section:ViscousInstabilities}, viscous instabilities is characterised by the formation of flow patterns, called viscous fingers. The geometrical evolution of these patterns provides good indications on capturing and tracking any instabilities in the flow front. Figure \ref{fig:3DChannel_PermMesh} shows the initial permeability distribution for the simulations described in this section. The surface on the left hand side of the domain is the area where fluid $1$ is injected with a uniform velocity of 2.5 cm.s$^{-1}$. No-slip conditions are applied everywhere else (\ie no flux across the boundaries) except at the right hand side of the domain where the fluid is allowed to leave the domain. Initially, the domain was partially saturated with non-wetting fluid, \ie $S_{nw}(t=0) = 0.80$, which was displaced by a pure (wetting) fluid. The domain was designed to capture preferential flow pathways, as shown in Fig.~\ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.

\medskip
Flow pathways are clearly indicated in Fig.~\ref{fig:3DChannel_planes}, where fluid $1$ Darcy velocity vectors overlapped with fluid saturation are shown. Fluid displacement occurs mainly in regions of relatively larger permeability, leaving other parts of the domain (\ie regions with low permeability distribution) with little or no momentum, indicating that at these regions the non-wetting fluid (\ie fluid 2) are not effectively displaced.

>>>>

In this section, 3D simulations are introduced to investigate preferential flow pathways and finger growth through semi-pervious and impervious layers. In large scale geological formations preferential flow may allow fast contaminant transport (therefore creating significant consequences for groundwater quality) and low recovery rate (in hydrocarbon production fields).

As outlined in Section \ref{section:ViscousInstabilities}, viscous instabilities are characterised by spontaneous formation and growth of fluid elongations, called viscous fingers. The geometrical evolution of these elongations provides good indication of instabilities at the fluid front interface. Figure \ref{fig:3DChannel_PermMesh} shows initial permeability distribution for the simulations described in this section. The surface on the left-hand side of the domain is the area where fluid $1$ is injected with a uniform velocity of 2.5 cm.s$^{-1}$. No-flux boundary conditions are applied to all borders, except the right-hand side (inlet) and the left-hand side (outflow). Initially, the domain was partially saturated with non-wetting fluid, \ie $S_{nw}(t=0)=0.80$, which was displaced by a pure (wetting) fluid. The domain (with 368513 \PNDG[1]{1} tetrahedral element-pairs) was designed to capture preferential flow pathways, \ie regions with larger permeability (Fig.~\ref{fig:3DChannel_PermMesh}) as demonstrated in Fig.~\ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.

Flow pathways are clearly indicated in Fig.~\ref{fig:3DChannel_planes}, where Darcy velocity vectors overlapped with wetting fluid saturation are shown. Fluid displacement occurs mainly in regions of relatively larger permeability, leaving other parts of the domain (\ie regions with low permeability distribution) with little or no momentum, indicating that at these regions the non-wetting fluid (\ie fluid 2) are not effectively displaced.
